---
title: "ML data details"
author: Sid (25 Nov 2019)
output: 
  html_document:
    code_folding: show
    number_sections: true
    toc: true
    toc_float: true
    theme: flatly
  df_print: paged
---

```{r, include=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE, out.width="55%", fig.align = "center")
```

```{r, include=FALSE}
options(stringsAsFactors=F)
Sys.setenv(http_proxy="http://internetproxy:8080")
library(tidyverse)
```


# ml_data_rf.rds  
This is my ML table for random forest with all the features I have quantified. 


# ml_data_sequnce.rds  
This data frame has the DNA sequences extracted to be used for the convolutional neural nets.  

I used the following code to extract the DNA sequences from genomic coordinates.  

```{r, eval=FALSE, include=TRUE}
library(tidyverse)
library(BSgenome)
library(BSgenome.Hsapiens.UCSC.hg38.masked)
library(GenomicRanges)


ml_table = readRDS(file = "ml_data_rf.rds") %>% 
  dplyr::select(class) %>% 
  rownames_to_column("id") %>% 
  tidyr::separate(id, c("gene", "seqnames", "start", "end", "strand", "tmp"), sep=":", remove=FALSE)
  

# converting to Granges object
ml_table_gr = makeGRangesFromDataFrame(ml_table, keep.extra.columns = TRUE)

# adding "chr" in front of seqnames
newStyle <- mapSeqlevels(seqlevels(ml_table_gr), "UCSC")
ml_table_gr <- renameSeqlevels(ml_table_gr, newStyle)

# Extract sequences using the package BSgenome
data = data.frame(
  id = ml_table$id,
  class = ml_table$class,
  X5utr = BSgenome::getSeq(Hsapiens, ml_table_gr, as.character = TRUE)
)

#save data
saveRDS(data, file = "ml_data_sequence.rds")
```


The "class" column is a factor with 6 levels. This would help you if you try to do a multi-class classification  

```{r}
table <- readRDS("ml_data_sequence.rds")
table(table$class)
```
  
The class "UTR" is our positive class (3'UTR). For tarining the random forest classifier, I use the following code to convert it into a binary class classification  

```{r}
table$class = ifelse(table$class %in% "UTR", "UTR", "Non_3_UTR")  %>% as.factor() %>% as.integer()
table$class <- table$class - 1 #0 = No3UTR 1 = 3UTR
table(table$class)
```

```{r}
PositiveSequences = table[table$class == 1,]
NegativeSequences = table[table$class == 0,]
```

```{r}
write.csv(PositiveSequences,file="../recognition-regulatory-sequences-thesis/Data/UTR/PositiveUTRSequences_untreated_sid", row.names = FALSE, quote = FALSE)
write.csv(NegativeSequences,file="../recognition-regulatory-sequences-thesis/Data/UTR/NegativeUTRSequences_untreated_sid", row.names = FALSE, quote = FALSE)
```



