---
title: "Training the network of the UTR experiment"
author: "Jaime Martínez Legaz"
output:
  html_document:
    df_print: paged
---

## DUDAS

- ¿No se obtiene el mismo resultado independientemente del caso?

## Loading the libraries

Loading all required libraries

```{r, eval=FALSE, echo=FALSE}
library(tensorflow)

with(tf$device("/gpu:0"), {

})

```

```{r}
library(SetupSequences)
library(keras)
library(caret)
library(stringi)
```

## Obtaining the sequences

We will create some simple sequences, very easy to differentiate. They will consist of 200 random characters (only T, C and G) and 50 Cs, either at the start of the sequence (positive), or at the end (negative). This way, we will have some easy to differentiate sequences, and we will be able to check if our LRP works as intended.

```{r}

PositiveSequences1 <- stri_rand_strings(500, 100, pattern = "[TGA]")
PositiveSequences2 <- stri_rand_strings(500, 100, pattern = "[TGA]")
NegativeSequences1 <- stri_rand_strings(500, 100, pattern = "[TGA]")
NegativeSequences2 <- stri_rand_strings(500, 100, pattern = "[TGA]")

Cseq <- paste(rep("C",50), collapse="")
Dseq <- paste(rep("X",50), collapse="")

PositiveSequences <- paste0(PositiveSequences1,PositiveSequences2,Cseq)
NegativeSequences <- paste0(NegativeSequences1,NegativeSequences2,Dseq)

```

## Treating the sequences

We start with the treatment of sequences. We only have to one-hot encode them.
For that, we need some functions from the custom package developed for this project.

```{r}
# to_onehot:   TGCCAA --> 00010010010010001000    (A = 1000, C = 0100, G = 0010, T = 0001)
PosSequences <- to_onehot(PositiveSequences)
NegSequences <- to_onehot(NegativeSequences)

PosSequences[1]  # The positive sequence, one-hot encoded
NegSequences[1]  # The negative sequence, one-hot encoded

# Validation
test1 = all(sapply(list(min(nchar(PosSequences)),max(nchar(PosSequences)), mean(nchar(PosSequences))), function(x) x == 1250))
test2 = all(sapply(list(min(nchar(NegSequences)),max(nchar(NegSequences)), mean(nchar(NegSequences))), function(x) x == 1250))
if (test1) print ("Correct 1") else print ("ERROR 1")
if (test2) print ("Correct 2") else print ("ERROR 2")
```

```{r}
# Final check: Since some of the sequences might be corrupt, we will delete the ones that are corrupt, if they exist
# We can know if a sequence is corrupt by looking for the W character. When one-hot encoding, we encoded everything that was not an A, T, C or G with a "W"

if (any(grepl("W",PosSequences))){
  print("Found 'W'")
  PosSequences <- PosSequences[-which(grepl("W",PosSequences)),]
}
  
if (any(grepl("W",NegSequences))){
  print("Found 'W'")
  NegSequences <- NegSequences[-which(grepl("W",NegSequences)),]
}
```

## Training the network
First we have to recover the treated data.

Now we face an issue with our sequences. They are 1250 characters long strings, but R sees them as indivisible elements. We have to transform them before working with them, turning each 1250 characters long sequence into 1250 sequences of only one character.

```{r}
posNum <- length(PosSequences) # Number of positive sequences
SeqUTR <- append(PosSequences,NegSequences)

old_seq = SeqUTR[1] # For validation tests

n2 <- nchar(SeqUTR[1]) -1 # Number of divisions to make

secuenciasOH <- sapply(1 + 1*0:n2, function(z) substr(SeqUTR, z, z))  # Obtaining the split characters
df <- data.frame(secuenciasOH, "Value" = 0) # Saving them into a dataframe
indx <- sapply(df, is.factor) 
df[indx] <- lapply(df[indx], function(x) as.character(x)) # Factor --> char conversion

df[1:posNum,]$Value <- 1 # We have the value of the positive sequences to positive

table(df$Value)

# Validation tests
test1 = paste(unlist(secuenciasOH[1,]),sep="", collapse="") == old_seq; if (test1) print ("Correct 1") else print ("ERROR 1")
test2 = table(df$Value)["0"] == length(NegSequences); if (test2) print ("Correct 2") else print ("ERROR 2")
test3 = table(df$Value)["1"] == length(PosSequences); if (test3) print ("Correct 3") else print ("ERROR 3")
test4 = all(length(PosSequences) + length(NegSequences) == length(SeqUTR), length(SeqUTR) == nrow(df)); if (test4) print ("Correct 4") else print ("ERROR 4")
```

Now we have to divide our data into Training and Test for our model training.

```{r}

output <- c("Value")
set.seed(1234)

partition <- createDataPartition(df[[output]],
                                     p = 0.8,
                                     list = FALSE,
                                     times = 1)
trains <- df[partition,]
tests <- df[-partition,]


```

Once divided, we have to adapt the data to the format the model expects.

```{r}

train.x <- data.matrix(trains[,1:1250])
train.y <- data.matrix(trains[,1251])
  
test.x <- data.matrix(tests[,1:1250])
test.y <- data.matrix(tests[,1251])

train.x <- array_reshape(train.x, c(nrow(trains),1250,1))
test.x <- array_reshape(test.x, c(nrow(tests),1250,1))
 
```

Now it is time to build our model. It follows the specifications mentioned in the thesis report. 

```{r}
  batch_size <- 125
  epochs <- 8 
  input_shape <- c(1250,1)
  learn_rate = 0.0001
  
  modelConvolu <- keras_model_sequential()
  modelConvolu %>% 
    layer_conv_1d(filters = 48, kernel_size = 75, activation = "relu", input_shape = input_shape)%>%
    layer_flatten() %>%
    layer_dense(units = 50, activation = 'relu') %>% 
    layer_dense(units = 1, activation="sigmoid") %>%
    
    summary(model)
  
  modelConvolu %>% compile(
    loss = loss_binary_crossentropy,
    optimizer = optimizer_nadam(lr = learn_rate),
    metrics = c('accuracy')
  )
```

And now we train that model with the data obtained. After each training epoch, we will obtain certain statistics related to the sensitivity and specificity of the model, since they are not metrics that Keras can return from the model.

```{r}
config = tensorflow::tf$ConfigProto(gpu_options = list(allow_growth = TRUE))
sess = tensorflow::tf$Session(config = config)
keras::k_set_session(session = sess)
```

```{r}
#with(tf$device("/device:GPU:0"), {

  
sens <- NULL
spec <- NULL
history <- NULL
 
  
modelConvolu <- keras_model_sequential()
  modelConvolu %>% 
    layer_conv_1d(filters = 48, kernel_size = 75, activation = "relu", input_shape = input_shape)%>%
    layer_flatten() %>%
    layer_dense(units = 50, activation = 'relu') %>% 
    layer_dense(units = 1, activation="sigmoid") 
  
  modelConvolu %>% compile(
    loss = loss_binary_crossentropy,
    optimizer = optimizer_nadam(lr = learn_rate),
    metrics = c('accuracy')
  )
  
    for (epoch in 1:epochs){
    historial <- modelConvolu %>% fit(
      x = train.x,
      y = train.y,
      epochs = 1,  
      batch_size = batch_size, 
      validation_data = list(test.x, test.y),
      verbose = 2)
    
    history <- c(history, historial)
  }

```

First, we have to obtain the data needed, for example the weights. For that, we need to access its layers 

```{r}
lrp <- function(model, dataframe, indexSeq){
  print(paste0("Sequence number ",indexSeq))
  layers <- model$layers
  input <- as.matrix(as.numeric(dataframe[indexSeq,1:1250])) # posNum+10
  
  ## STEP 1: CREATE DATA STRUCTURES
  activs <- list()
  weights <- list()
  biases <- list()

  for (i in 1:length(layers)){
    lay <- layers[[i]]
    layer_name <- lay$name
    
    intermediate_layer_model <- keras_model(inputs = modelConvolu$input,
                                      outputs = get_layer(modelConvolu, layer_name)$output)
    intermediate_output <- predict(intermediate_layer_model, array_reshape(input, c(1,1250,1)))
    
    if (length(lay$weights) > 0){ # If not a flatten layer
      activs[[i]] <- intermediate_output
      weights[[i]] <- lay$get_weights()[[1]]
      biases[[i]] <- lay$get_weights()[[2]]
    }
  }
  
  ## STEP 2: OUTPUT LAYER
  
  Rk <- list()
  Rk[["output"]] <- as.numeric(activs[[length(layers)]])
  print(Rk[["output"]])
  
  ## STEP 3: DENSE LAYER
  
  pos_weights <- weights[[4]]
  for (i in 1:length(weights[[4]]))
    pos_weights[i] <- max(weights[[4]][i],0) # Not using the apply family functions so it does not alter the shape
  
##  if (min(pos_weights) == 0) print("CORRECT 1") else print ("ERROR 1")
  
  denom = 0
  
  for (i in 1:length(activs[[3]])) { # Calculating the denominator
    value <- activs[[3]][i] * pos_weights[i]
    denom <- denom + value
  }
  
  for (i in 1:length(activs[[3]])) { # Calculating each numerator and, with it, the relevance of each neuron of the dense layer
    numer <- activs[[3]][i] * pos_weights[i]
    if (denom != 0) {
        division <- numer/denom
      } else {
        division <- 0
      }
    Rk[["dense"]][i] <- division * Rk[["output"]][1]
  }
  
  ## STEP 4: CONVOLUTIONAL LAYER
  
  pos_weights <- weights[[3]]
  for (i in 1:length(weights[[3]]))
    pos_weights[i] <- max(weights[[3]][i],0) # Not using the apply family functions so it does not alter the shape
  
##  if (min(pos_weights) == 0) print("CORRECT 1") else print ("ERROR 1")
  
  for (i in 1:length(activs[[1]]))
    Rk[["conv"]][i] = 0
  
  for (k in 1:length(Rk[["dense"]])) { # For each neuron in the layer analyzed before we have to repeat the main cycle
    denom <- 0
    
    # pos_weights[j,k]
  #  individual_weights = length(pos_weights) / length(Rk[["dense"]])
    
    for (j in 1:length(activs[[1]])) { # For each neuron in the actual layer, calculate denominator
      value <- activs[[1]][j] * pos_weights[j,k]
      denom <- denom + value
    }
    
    for (j in 1:length(activs[[1]])) {# For each neuron in the actual layer, calculate numerator and relevance j<-k
      numer <- activs[[1]][j] * pos_weights[j,k]
      if (denom != 0) {
        division <- numer/denom
      } else {
        division <- 0
      }
      Rk[["conv"]][j] <- Rk[["conv"]][j] + (division * Rk[["dense"]][k])
    }
  }
  
  ## STEP 5: INPUT LAYER
  
  pos_weights <- weights[[1]]

  for (i in 1:length(weights[[1]]))
    pos_weights[i] <- max(weights[[1]][i],0) # Not using the apply family functions so it does not alter the shape
  
##  if (min(pos_weights) == 0) print("CORRECT 1") else print ("ERROR 1")
  
  
  used <- list()
  for (i in 1:1250){
    Rk[["input"]][i] = 0
    used[["input"]][i] = 0
  }
  
  for (k in 1:length(activs[[1]])){ # For each one of the convolution cells
  #  if (Rk[["conv"]][k] == 0)
   #   next
    ## First we have to obtain the list of input nodes that contribute to this value. We know that there are as many nodes as the size of the filters:
    num_neurons <- length(weights[[1]][,,1])
    
    ## Now we have to know which are these neurons. For that reason, we have to obtain the position of this cell relative to the convolution filter
    # Since a%%b returns a value in [0,b-1] and we want a value in [1,b], we have to do it this way:
    cells_per_kernel <- length(activs[[1]][,,1])
    k_relative <- ((k-1)%%cells_per_kernel)+1
    # And now we obtain the list of input values that affected this particular cell
    j_neurons <- k_relative:(k_relative + num_neurons - 1)
    
    ## To obtain the weights, we have to know to which kernel this cell belongs to
    kernel <- ceiling(k/cells_per_kernel)
    
    ## And now we can obtain said weights
    kernel_weights <- pos_weights[,,kernel]
    
    ## Now we can calculate the denominator for this filter, following the formula
    denom <- sum(input[j_neurons] * kernel_weights)
    
    ## And, for each j neuron, we calculate its numerator, and add its value to its relevance
    for (j in 1:length(j_neurons)){
      numer <- input[j_neurons[j]] * kernel_weights[j] * Rk[["conv"]][k]
      if (denom != 0) {
        division <- numer/denom
      } else {
        division <- 0
      }
      Rk[["input"]][j_neurons[j]] = Rk[["input"]][j_neurons[j]] + division
      used[["input"]][j_neurons[j]] = used[["input"]][j_neurons[j]] + 1
    }
  }
  
  ## STEP 6: RETURNING RESULT
  
 return(Rk[["input"]] / used[["input"]])
#return(Rk[["input"]])
    
}
```

```{r, eval=FALSE}

###### DEFINIR FUNCIÓN QUE ENGLOBE TODO ESTO. APLICAR LA FUNCIÓN A TODAS LAS SECUENCIAS, SUMANDO LOS RESULTADOS. PLOTTEARLOS.

layers <- modelConvolu$layers
input <- as.matrix(as.numeric(df[1,1:1250]))

activs <- list()
weights <- list()
biases <- list()

for (i in 1:length(layers)){
  lay <- layers[[i]]
  layer_name <- lay$name
  
  intermediate_layer_model <- keras_model(inputs = modelConvolu$input,
                                    outputs = get_layer(modelConvolu, layer_name)$output)
  intermediate_output <- predict(intermediate_layer_model, array_reshape(input, c(1,1250,1)))
  
  if (length(lay$weights) > 0){ # If not a flatten layer
    activs[[i]] <- intermediate_output
    weights[[i]] <- lay$get_weights()[[1]]
    biases[[i]] <- lay$get_weights()[[2]]
  }
}

```

Now that we have the information required, we can start propagating the relevance.

```{r, LRP-outpu, eval=FALSE}

### OUTPUT LAYER
# Obtain the output value of the output node = Rk
Rk <- list()
Rk[["output"]] <- as.numeric(activs[[length(layers)]])

```

```{r, LRP-dense, eval=FALSE}

### DENSE LAYER
# For each neuron, apply the formula to obtain its relevance
# pos_weights <- sapply(weights[[3]], function(x) max(x,0)) # Obtains the positive parts of the weights

pos_weights <- weights[[4]]
for (i in 1:length(weights[[4]]))
  pos_weights[i] <- max(weights[[4]][i],0) # Not using the apply family functions so it does not alter the shape

if (min(pos_weights) == 0) print("CORRECT 1") else print ("ERROR 1")

denom = 0

for (i in 1:length(activs[[3]])) { # Calculating the denominator
  value <- activs[[3]][i] * pos_weights[i]
  denom <- denom + value
}

for (i in 1:length(activs[[3]])) { # Calculating each numerator and, with it, the relevance of each neuron of the dense layer
  numer <- activs[[3]][i] * pos_weights[i]
  if (denom != 0) {
      division <- numer/denom
    } else {
      division <- 0
    }
  Rk[["dense"]][i] <- division * Rk[["output"]][1]
}

```
```{r, LRP-conv, eval=FALSE}
pos_weights <- weights[[3]]
for (i in 1:length(weights[[3]]))
  pos_weights[i] <- max(weights[[3]][i],0) # Not using the apply family functions so it does not alter the shape

if (min(pos_weights) == 0) print("CORRECT 1") else print ("ERROR 1")

for (i in 1:length(activs[[1]]))
  Rk[["conv"]][i] = 0

for (k in 1:length(Rk[["dense"]])) { # For each neuron in the layer analyzed before we have to repeat the main cycle
  denom <- 0
  
  # pos_weights[j,k]
#  individual_weights = length(pos_weights) / length(Rk[["dense"]])
  
  for (j in 1:length(activs[[1]])) { # For each neuron in the actual layer, calculate denominator
    value <- activs[[1]][j] * pos_weights[j,k]
    denom <- denom + value
  }
  
  for (j in 1:length(activs[[1]])) {# For each neuron in the actual layer, calculate numerator and relevance j<-k
    numer <- activs[[1]][j] * pos_weights[j,k]
    if (denom != 0) {
      division <- numer/denom
    } else {
      division <- 0
    }
    Rk[["conv"]][j] <- Rk[["conv"]][j] + (division * Rk[["dense"]][k])
  }
}

```
```{r, Hipotesis, eval = FALSE, echo = FALSE}

## PRUEBA DE FUNCIONAMIENTO DE CONV1D
## Hipótesis: El valor #2 del resultado (que sería la activación #2 de la capa 1) es igual a multiplicar los valores 2:76 de la secuencia test.x[[1]][1,,] por los pesos [1:75,,1]
result <- activs[[1]][2]
valores <- test.x[[1]][1,2:76,]
pesos <- weights[[1]][,,1]

sum = 0
for (i in 1:75) {
  sum = sum + (valores[i]*pesos[i])
}

suma2 = sum + biases[[1]][1]
## Hipótesis CONFIRMADA
# El valor #i de la capa post-convolución es igual a multiplicar los valores i:(i+75) de la secuencia por los pesos [1:75,,n] (Para el filtro n, el cual sabremos con n = i mod 75)

```

```{r, LRP-input_old, eval=FALSE, echo=FALSE}

pos_weights <- weights[[1]]
for (i in 1:length(weights[[1]]))
  pos_weights[i] <- max(weights[[1]][i],0) # Not using the apply family functions so it does not alter the shape

if (min(pos_weights) == 0) print("CORRECT 1") else print ("ERROR 1")

for (i in 1:1250)
  Rk[["input"]][i] = 0

for (j in 1:length(test.x[[1]][1,,])){
  affected_indexes <- seq((j-75),j) # Index of the convolution cells affected by this value
  affected_indexes = affected_indexes[affected_indexes > 0 & affected_indexes <= length(activs[[1]][,,1])] # We remove the invalid values: Those lesser than zero, and those greater than the length of the result
  for (i in 1:length(affected_indexes)){
    k <- affected_indexes[i] # The "k"" neuron of the post-convolution layer
    jneurons <- k:(k+75) # The "j" neurons of the pre-convolution layer that are used to calculate "k"'s activation value 
    
    denom <- 0
    for (n in 1:length(jneurons)){
      j <- jneurons[n] # We grab the nth neuron of this group
      value <- input[j] * 
      
    }
  }
}

```

```{r, LRP-input, eval=FALSE}
pos_weights <- weights[[1]]

for (i in 1:length(weights[[1]]))
  pos_weights[i] <- max(weights[[1]][i],0) # Not using the apply family functions so it does not alter the shape

if (min(pos_weights) == 0) print("CORRECT 1") else print ("ERROR 1")


used <- list()
for (i in 1:1250){
  Rk[["input"]][i] = 0
  used[["input"]][i] = 0
}

## PLAN
## For each one of the 56448 convolution cells,
## - We obtain the list of input neurons (j) that contributed to its activation
## - We obtain the filter it belongs to
## - We obtain its 75 weights
## - We calculate the denominator for that filter
## - For each one of the 75 j neurons
## -- We add the result to the activation value of that neuron

for (k in 1:length(activs[[1]])){ # For each one of the convolution cells
#  if (Rk[["conv"]][k] == 0)
 #   next
  ## First we have to obtain the list of input nodes that contribute to this value. We know that there are as many nodes as the size of the filters:
  num_neurons <- length(weights[[1]][,,1])
  
  ## Now we have to know which are these neurons. For that reason, we have to obtain the position of this cell relative to the convolution filter
  # Since a%%b returns a value in [0,b-1] and we want a value in [1,b], we have to do it this way:
  cells_per_kernel <- length(activs[[1]][,,1])
  k_relative <- ((k-1)%%cells_per_kernel)+1
  # And now we obtain the list of input values that affected this particular cell
  j_neurons <- k_relative:(k_relative + num_neurons - 1)
  
  ## To obtain the weights, we have to know to which kernel this cell belongs to
  kernel <- ceiling(k/cells_per_kernel)
  
  ## And now we can obtain said weights
  kernel_weights <- pos_weights[,,kernel]
  
  ## Now we can calculate the denominator for this filter, following the formula
  denom <- sum(input[j_neurons] * kernel_weights)
  
  ## And, for each j neuron, we calculate its numerator, and add its value to its relevance
  for (j in 1:length(j_neurons)){
    numer <- input[j_neurons[j]] * kernel_weights[j] * Rk[["conv"]][k]
    if (denom != 0) {
      division <- numer/denom
    } else {
      division <- 0
    }
    Rk[["input"]][j_neurons[j]] = Rk[["input"]][j_neurons[j]] + division
    used[["input"]][j_neurons[j]] = used[["input"]][j_neurons[j]] + 1
  }
}

Rk[["input"]] <- Rk[["input"]] / used[["input"]]

```

We obtain the relevance to each case, and then we add them.

```{r}

posCases <- which(df$Value == 1)
negCases <- which(df$Value == 0)

allRelevances <- list()
allRelevances[["pos"]] <- list()
allRelevances[["neg"]] <- list()


for (i in posCases[1:20]){
  allRelevances[["pos"]] <- append(allRelevances[["pos"]], list(lrp(modelConvolu, df, i)))
}

for (i in negCases[1:20]){
  allRelevances[["neg"]] <- append(allRelevances[["neg"]], list(lrp(modelConvolu, df, i)))
}

reli <- unlist(allRelevances[["neg"]][1])
for (i in 2:20){
  reli <- reli + unlist(allRelevances[["neg"]][i])
}
plot(reli,type="l")


```

```{r, Obtain_relevances_like_script}

relevancesP <- as.data.frame(PosSequences, stringsAsFactors = FALSE)
relevancesN <- as.data.frame(NegSequences, stringsAsFactors = FALSE)

posS <- df[df$Value == 1,] # Quizá cambiar al formato del BasicExample, con el which
negS <- df[df$Value == 0,]

for (i in 1:250){
  relevancesP[[paste0("relevance",i)]] <- 0
  relevancesN[[paste0("relevance",i)]] <- 0
}

for (i in 1:30){
  relevances <- lrp(modelConvolu, posS, i)
  
  for (j in 1:250){
    k = 1 + (5*(j - 1))
    relevancesP[i,paste0("relevance", j)] <- relevances[k] + relevances[k + 1] + relevances[k + 2] + relevances[k + 3] + relevances[k + 4]
  }
}

for (i in 1:30){
  relevances <- lrp(modelConvolu, negS, i)
  
  for (j in 1:250){
    k = 1 + (5*(j - 1))
    relevancesN[i,paste0("relevance", j)] <- relevances[k] + relevances[k + 1] + relevances[k + 2] + relevances[k + 3] + relevances[k + 4]
  }
}
```

```{r}

# Normalized the values of a dataframe (max 1, min 0)

normalizeDF <- function(dataframe){
 mn <- min(dataframe)
 mx <- max(dataframe)
  return((dataframe - mn) / (mx - mn))
}

 
```
 
```{r}

pr <- relevancesP[2:251]
nr <- relevancesN[2:251]

```

```{r}
## Sin normalización (MAL)
p_rel <- colSums(pr)      
n_rel <- colSums(nr)

plot(colSums(pr[1,]), type="l")
for (i in 2:(nrow(pr))) {
  lines(colSums(pr[i,]))
}
plot(colSums(pr),type="l")

plot(colSums(nr[1,]), type="l")
for (i in 2:(nrow(nr))) {
  lines(colSums(nr[i,]))
}

plot(colSums(nr),type="l")
```

```{r}
## Aplicando normalización

pr_norm <- normalizeDF(pr)
nr_norm <- normalizeDF(nr)

plot(colSums(pr_norm[1,]), type="l", ylim=c(0,1))
for (i in 2:(nrow(pr_norm))) {
  lines(colSums(pr_norm[i,]))
}
plot(colSums(pr_norm),type="l")

plot(colSums(nr_norm[1,]), type="l", ylim=c(0,1))
for (i in 2:(nrow(nr_norm))) {
  lines(colSums(nr_norm[i,]))
}
plot(colSums(nr_norm),type="l")
```


