---
title: "Treatment of sequences for the Promoters vs UTR experiment"
author: "Jaime Martínez Legaz"
output: html_document
---

## Initial setup

Loading all required libraries

```{r}
library(SetupSequences)
```

## Treating the sequences

First, we recover the sequences we obtained with ensembl.

```{r}
PromSequences <- read.csv("Data/PromvsUTR/PromoterSequences_untreated")
UTRSequences <- read.csv("Data/PromvsUTR/UTRSequences_untreated")
```

The format is a bit different from what we look for. We have to change the type of the columns, since we are working with characters, not factors.

```{r}

PromSequences$hgnc_symbol <- as.character(PromSequences$hgnc_symbol)
PromSequences$coding_gene_flank <- as.character(PromSequences$coding_gene_flank)

UTRSequences$hgnc_symbol <- as.character(UTRSequences$hgnc_symbol)
UTRSequences$X5utr <- as.character(UTRSequences$X5utr)

```

Next we start with the treatment of sequences. We have to:

- Reverse them
- Make sure all sequences are 250 characters long (shorter UTR sequences have already been discarded, though)
- One-hot encode them

For that, we need some functions from the custom package developed for this project.

```{r}
# Example of sequence: AACCGT
PromSequences$coding_gene_flank[1]  # Example of a Promoter sequence
UTRSequences$X5utr[1]  # Example of a UTR sequence

# strReverse:  AACCGT --> TGCCAA
PromSequences$coding_gene_flank <- strReverse(PromSequences$coding_gene_flank)
UTRSequences$X5utr <- strReverse(UTRSequences$X5utr)

PromSequences$coding_gene_flank[1]  # The promoter sequence, once reversed
UTRSequences$X5utr[1]  # The UTR sequence, once reversed

# padding_sequences: Inserts padding characters in sequences shorter than 250 characters, and trims sequences longer than 250 characters
UTRSequences$X5utr <- padding_sequences(UTRSequences$X5utr)  

UTRSequences$X5utr[1]  # The UTR sequence, trimmed to be 250 characters long
nchar(UTRSequences$X5utr[1])  # Length of said UTR sequence

# to_onehot:   TGCCAA --> 0001000100010001000010000    (A = 10000, C = 01000, G = 00100, T = 00010, X = 00001)
PromSequences$coding_gene_flank <- to_onehot(PromSequences$coding_gene_flank)
UTRSequences$X5utr <- to_onehot(UTRSequences$X5utr)

PromSequences$coding_gene_flank[1]  # The promoter sequence, once reversed and one-hot encoded
UTRSequences$X5utr[1]  # The UTR sequence, once reversed, trimmed and one-hot encoded

# Final check: Since some of the sequences might be corrupt, we will delete the ones that are corrupt, if they exist
# We can know if a sequence is corrupt by looking for the W character. When one-hot encoding, we encoded everything that was not an A, T, C or G with a "W"

if (any(grepl("W",UTRSequences$X5utr)))
  UTRSequences <- UTRSequences[-which(grepl("W",UTRSequences$X5utr)),]

if (any(grepl("W",PromSequences$coding_gene_flank)))
  PromSequences <- PromSequences[-which(grepl("W",PromSequences$coding_gene_flank)),]
```

Once treated, we can save them in a file for later use.

```{r}
write.csv(PromSequences,file="Data/PromvsUTR/PromoterSequences_treated", row.names = FALSE, quote = FALSE)
write.csv(UTRSequences,file="Data/PromvsUTR/UTRSequences_treated", row.names = FALSE, quote = FALSE)

write.table(PromSequences$coding_gene_flank,file="Data/PromvsUTR/PromoterSequences_sequences_treated", row.names = FALSE, quote = FALSE, col.names = FALSE)
write.table(UTRSequences$X5utr,file="Data/PromvsUTR/UTRSequences_sequences_treated", row.names = FALSE, quote = FALSE, col.names = FALSE)
```

