---
title: "Treatment of sequences for the Promoters experiment"
author: "Jaime Martínez Legaz"
output: html_document
---

## Initial setup

Loading all required libraries

```{r}
library(SetupSequences)
```

## Treating the sequences

First, we recover the sequences we obtained with ensembl.

```{r}
PosSequences <- read.csv("Data/UTR/PositiveUTRSequences_untreated")
NegSequences <- read.csv("Data/UTR/NegativeUTRSequences_untreated")
```

The format is a bit different from what we look for. We have to change the type of the columns, since we are working with characters, not factors.

```{r}

PosSequences$ensembl_gene_id <- as.character(PosSequences$ensembl_gene_id)
PosSequences$X5utr <- as.character(PosSequences$X5utr)

NegSequences$ensembl_gene_id <- as.character(NegSequences$ensembl_gene_id)
NegSequences$X5utr <- as.character(NegSequences$X5utr)

```

Next we start with the treatment of sequences. We have to:

- Reverse them
- Make sure all sequences are 250 characters long
- One-hot encode them

For that, we need some functions from the custom package developed for this project.

```{r}
# Example of sequence: AACCGT
PosSequences$X5utr[1]  # Example of a positive sequence
NegSequences$X5utr[1]  # Example of a negative sequence

# strReverse:  AACCGT --> TGCCAA
PosSequences$X5utr <- strReverse(PosSequences$X5utr)
NegSequences$X5utr <- strReverse(NegSequences$X5utr)

PosSequences$X5utr[1]  # The positive sequence, once reversed
NegSequences$X5utr[1]  # The negative sequence, once reversed

# padding_sequences: Inserts padding characters ("X") in sequences shorter than 250 characters, and trims sequences longer than 250 characters
PosSequences$X5utr <- padding_sequences(PosSequences$X5utr)
NegSequences$X5utr <- padding_sequences(NegSequences$X5utr)

PosSequences$X5utr[1]  # The positive sequence, with some characters added so it can be 250 characters long
NegSequences$X5utr[1]  # The negative sequence, trimmed so it becomes 250 characters long

# to_onehot:   TGCCAA --> 00010010010010001000    (A = 1000, C = 0100, G = 0010, T = 0001)
PosSequences$X5utr <- to_onehot(PosSequences$X5utr)
NegSequences$X5utr <- to_onehot(NegSequences$X5utr)

PosSequences$X5utr[1]  # The positive sequence, reversed, padded and one-hot encoded
NegSequences$X5utr[1]  # The negative sequence, reversed, trimmed and one-hot encoded

# Final check: Since some of the sequences might be corrupt, we will delete the ones that are corrupt, if they exist
# We can know if a sequence is corrupt by looking for the W character. When one-hot encoding, we encoded everything that was not an A, T, C or G with a "W"

if (any(grepl("W",PosSequences$X5utr)))
  PosSequences <- PosSequences[-which(grepl("W",PosSequences$X5utr)),]

if (any(grepl("W",NegSequences$X5utr)))
  NegSequences <- NegSequences[-which(grepl("W",NegSequences$X5utr)),]
```

Once treated, we can save them in a file for later use.

```{r}
write.csv(PosSequences,file="Data/UTR/PositiveUTRSequences_treated", row.names = FALSE, quote = FALSE)
write.csv(NegSequences,file="Data/UTR/NegativeUTRSequences_treated", row.names = FALSE, quote = FALSE)

write.table(PosSequences$X5utr,file="Data/UTR/PositiveUTRSequences_sequences_treated", row.names = FALSE, quote = FALSE, col.names = FALSE)
write.table(NegSequences$X5utr,file="Data/UTR/NegativeUTRSequences_sequences_treated", row.names = FALSE, quote = FALSE, col.names = FALSE)

seqUTR <- append(PosSequences$X5utr,NegSequences$X5utr)
fileCon<-file("Data/UTR/SeqUTR")
write.table(seqUTR,file = fileCon, quote=FALSE, row.names = FALSE, col.names = FALSE)
```

