---
title: "Treatment of sequences for the Promoters experiment"
author: "Jaime Martínez Legaz"
output: html_document
---

## Initial setup

Loading all required libraries

```{r}
library(SetupSequences)
```

## Treating the sequences

First, we recover the sequences we obtained with ensembl.

```{r}
PosSequences <- read.csv("Data/UTR/PositiveUTRSequences_untreated", stringsAsFactors = F)
NegSequences <- read.csv("Data/UTR/NegativeUTRSequences_untreated", stringsAsFactors = F)
```

The format is a bit different from what we look for. We have to change the type of the columns, since we are working with characters, not factors.

```{r}
# BORRABLE
PosSequences$ensembl_gene_id <- as.character(PosSequences$ensembl_gene_id)
PosSequences$X5utr <- as.character(PosSequences$X5utr)

NegSequences$ensembl_gene_id <- as.character(NegSequences$ensembl_gene_id)
NegSequences$X5utr <- as.character(NegSequences$X5utr)

```

Now we have to get the direction of the strand of each gene, since it will decide whether we should or not reverse our sequences for better results. For that, we will use the following file, generated by the "UTR_strandDirection" script.

```{r, echo=FALSE}

strands <-  read.delim("Data/UTR/strands.tsv",
                      stringsAsFactors=F,header=T)

PosSequences$X5utr <- strReverse(PosSequences$X5utr)
NegSequences$X5utr <- strReverse(NegSequences$X5utr)

PosSequences$direction <- "?"
NegSequences$direction <- "?" 

for (i in 1:nrow(PosSequences)){
  print(i)
  gene <- PosSequences$hgnc_symbol[i]
  seq <- PosSequences$X5utr[i]
  
  for (j in 1:nrow(strands)){
    if ((strands$gene[j] == gene)&&(strands$utr[j] == seq))
      PosSequences$direction[i] <- strands$strand[j]
  }
  
#  if (any(grepl(paste0("^",gene,"$"),strands$gene) && (grepl(paste0("^",seq,"$"),strands$utr))))
#     PosSequences$direction[i] <- strands$strand[(strands$gene == gene) && (strands$utr == seq)]
}
  
for (i in 1:nrow(NegSequences)){
  print(i)
  gene <- NegSequences$hgnc_symbol[i]
  seq <- NegSequences$X5utr[i]
  
  for (j in 1:nrow(strands)){
    if ((strands$gene[j] == gene)&&(strands$utr[j] == seq))
      NegSequences$direction[i] <- strands$strand[j]
  }
  
  #if (any(grepl(paste0("^",gene,"$"),strands$gene) && (grepl(paste0("^",seq,"$"),strands$utr))))
  #   NegSequences$direction[i] <- strands$strand[strands$gene == gene]
}
```
```{r}
table(PosSequences$direction)
table(NegSequences$direction)
```

Next we start with the treatment of sequences. We have to:

- Reverse them (only if negative strand)
- Make sure all sequences are 250 characters long
- One-hot encode them

For that, we need some functions from the custom package developed for this project.

```{r}
# Example of sequence: AACCGT
PosSequences$X5utr[1]  # Example of a positive sequence
NegSequences$X5utr[1]  # Example of a negative sequence

# strReverse:  AACCGT --> TGCCAA
#PosSequences$X5utr <- strReverse(PosSequences$X5utr)
#NegSequences$X5utr <- strReverse(NegSequences$X5utr)

PosSequences$X5utr[1]  # The positive sequence, once reversed
NegSequences$X5utr[1]  # The negative sequence, once reversed

# padding_sequences: Inserts padding characters ("X") in sequences shorter than 250 characters, and trims sequences longer than 250 characters
PosSequences$X5utr <- padding_sequences_left(PosSequences$X5utr)
NegSequences$X5utr <- padding_sequences_left(NegSequences$X5utr)

PosSequences$X5utr[1]  # The positive sequence, with some characters added so it can be 250 characters long
NegSequences$X5utr[1]  # The negative sequence, trimmed so it becomes 250 characters long

# to_onehot:   TGCCAA --> 00010010010010001000    (A = 1000, C = 0100, G = 0010, T = 0001)
PosSequences$X5utr <- to_onehot(PosSequences$X5utr)
NegSequences$X5utr <- to_onehot(NegSequences$X5utr)

PosSequences$X5utr[1]  # The positive sequence, reversed, padded and one-hot encoded
NegSequences$X5utr[1]  # The negative sequence, reversed, trimmed and one-hot encoded

# Final check: Since some of the sequences might be corrupt, we will delete the ones that are corrupt, if they exist
# We can know if a sequence is corrupt by looking for the W character. When one-hot encoding, we encoded everything that was not an A, T, C or G with a "W"

if (any(grepl("W",PosSequences$X5utr)))
  PosSequences <- PosSequences[-which(grepl("W",PosSequences$X5utr)),]

if (any(grepl("W",NegSequences$X5utr)))
  NegSequences <- NegSequences[-which(grepl("W",NegSequences$X5utr)),]
```

Once treated, we can save them in a file for later use.

```{r}
write.csv(PosSequences,file="Data/UTR/PositiveUTRSequences_treated", row.names = FALSE, quote = FALSE)
write.csv(NegSequences,file="Data/UTR/NegativeUTRSequences_treated", row.names = FALSE, quote = FALSE)

write.table(PosSequences$X5utr,file="Data/UTR/PositiveUTRSequences_sequences_treated", row.names = FALSE, quote = FALSE, col.names = FALSE)
write.table(NegSequences$X5utr,file="Data/UTR/NegativeUTRSequences_sequences_treated", row.names = FALSE, quote = FALSE, col.names = FALSE)

seqUTR <- append(PosSequences$X5utr,NegSequences$X5utr)
fileCon<-file("Data/UTR/SeqUTR")
write.table(seqUTR,file = fileCon, quote=FALSE, row.names = FALSE, col.names = FALSE)
```

