---
title: "Treatment of sequences for the Promoters experiment"
author: "Jaime Martínez Legaz"
output: html_document
---

## Initial setup

Loading all required libraries

```{r}
library(SetupSequences)
```

## Treating the sequences

First, we recover the sequences we obtained with ensembl.

```{r}
PosSequences <- read.csv("Data/Promoters/PositivePromoterSequences_untreated")
NegSequences <- read.csv("Data/Promoters/NegativePromoterSequences_untreated")
```

The format is a bit different from what we look for. We have to change the type of the columns, since we are working with characters, not factors.

```{r}

PosSequences$hgnc_symbol <- as.character(PosSequences$hgnc_symbol)
PosSequences$coding_gene_flank <- as.character(PosSequences$coding_gene_flank)

NegSequences$hgnc_symbol <- as.character(NegSequences$hgnc_symbol)
NegSequences$coding_gene_flank <- as.character(NegSequences$coding_gene_flank)

```

Next we start with the treatment of sequences. We have to:

- Reverse them
- One-hot encode them

For that, we need some functions from the custom package developed for this project.

```{r}
# Example of sequence: AACCGT
PosSequences$coding_gene_flank[1]  # Example of a positive sequence
NegSequences$coding_gene_flank[1]  # Example of a negative sequence

# strReverse:  AACCGT --> TGCCAA
PosSequences$coding_gene_flank <- strReverse(PosSequences$coding_gene_flank)
NegSequences$coding_gene_flank <- strReverse(NegSequences$coding_gene_flank)

PosSequences$coding_gene_flank[1]  # The positive sequence, once reversed
NegSequences$coding_gene_flank[1]  # The negative sequence, once reversed

# to_onehot:   TGCCAA --> 0001000100010001000010000    (A = 10000, C = 01000, G = 00100, T = 00010, X = 00001)
PosSequences$coding_gene_flank <- padding_onehot(PosSequences$coding_gene_flank)
NegSequences$coding_gene_flank <- padding_onehot(NegSequences$coding_gene_flank)

PosSequences$coding_gene_flank[1]  # The positive sequence, once reversed and one-hot encoded
NegSequences$coding_gene_flank[1]  # The negative sequence, once reversed and one-hot encoded

# Final check: Since some of the sequences might be corrupt, we will delete the ones that are corrupt, if they exist
# We can know if a sequence is corrupt by looking for the W character. When one-hot encoding, we encoded everything that was not an A, T, C or G with a "W"

if (any(grepl("W",PosSequences$coding_gene_flank)))
  PosSequences <- PosSequences[-which(grepl("W",PosSequences$coding_gene_flank)),]

if (any(grepl("W",NegSequences$coding_gene_flank)))
  NegSequences <- NegSequences[-which(grepl("W",NegSequences$coding_gene_flank)),]
```

Once treated, we can save them in a file for later use.

```{r}
write.csv(PosSequences,file="Data/Promoters/PositivePromoterSequences_treated", row.names = FALSE, quote = FALSE)
write.csv(NegSequences,file="Data/Promoters/NegativePromoterSequences_treated", row.names = FALSE, quote = FALSE)

seqProm <- append(PosSequences$coding_gene_flank,NegSequences$coding_gene_flank)
fileCon<-file("Data/Promoters/SeqProm")
write.table(seqProm,file = fileCon, quote=FALSE, row.names = FALSE, col.names = FALSE)
```

