---
title: "Training the network of the UTR experiment"
author: "Jaime Martínez Legaz"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    toc_collapsed: true
    pdf_document: default
    toc_depth: 3
    number_sections: true
    theme: lumen
---


```{r, eval=FALSE, echo=FALSE}
library(tensorflow)

with(tf$device("/gpu:0"), {

})

```

```{r, include=FALSE}
library(biomaRt)
library(SetupSequences)
library(keras)
library(caret)
library(tidyverse)
library(BSgenome)
library(BSgenome.Hsapiens.UCSC.hg38.masked)
library(GenomicRanges)
```

# Obtaining the sequences

We have to obtain the sequences we will work with. This is the code that will achieve that.

```{r}
ml_table = readRDS(file = "../Data/ml_data_rf.rds") %>% 
  dplyr::select(class) %>% 
  rownames_to_column("id") %>% 
  tidyr::separate(id, c("gene", "seqnames", "start", "end", "strand", "tmp"), sep=":", remove=FALSE)
  

# converting to Granges object
ml_table_gr = makeGRangesFromDataFrame(ml_table, keep.extra.columns = TRUE)

# adding "chr" in front of seqnames
newStyle <- mapSeqlevels(seqlevels(ml_table_gr), "UCSC")
ml_table_gr <- renameSeqlevels(ml_table_gr, newStyle)

# Extract sequences using the package BSgenome
data = data.frame(
  id = ml_table$id,
  class = ml_table$class,
  sequence = BSgenome::getSeq(Hsapiens, ml_table_gr, as.character = TRUE) #as.character=TRUE does not work?
)
data$sequence <- as.character(data$sequence)
table(data$class)
```

Next step is grouping all the non 3'UTR sequences into a group, so that we can compare 3'UTR sequences with non 3'UTR sequences regardless of the subtype.

```{r}
old_data <- data # Needed for the valdation tests

data <- data[data$class == "ICE" | data$class == "UTR",]
data$detailclass = data$class
data$class = ifelse(data$class %in% "UTR", "UTR", "Non_3_UTR")  %>% as.factor() %>% as.integer()
data$class <- data$class - 1 #0 = No3UTR 1 = 3UTR
table(data$class)
```

```{r}
# Validation test
test1 = table(data$class)["0"] == nrow(data) - table(old_data$class)["UTR"]; if (test1 == FALSE) print ("ERROR 1")
test2 = table(data$class)["1"] == table(old_data$class)["UTR"]; if (test2 == FALSE) print ("ERROR 2")

```

We will separate these sequences into the "Positive" group (3'UTR) and the "Negative" group (non-3'UTR).

```{r}
PositiveSequences = data[data$class == 1,]
NegativeSequences = data[data$class == 0,]
```

# Studying different lengths of sequences

## With total numbers

```{r}

sizes <- seq(100,400)
num_total <- NULL
num_pos <- NULL
num_neg <- NULL

for (i in sizes){
  npos <- sum(nchar(PositiveSequences$sequence) > i)
  nneg <- sum(nchar(NegativeSequences$sequence) > i)
  ntot <- (sum(nchar(PositiveSequences$sequence) > i) + sum(nchar(NegativeSequences$sequence) > i))
  
  num_total <- c(num_total, ntot)
  num_pos <- c(num_pos, npos)
  num_neg <- c(num_neg, nneg)
}

df <- data.frame(sizes = sizes, npos <- num_pos, nneg <- num_neg, ntotal <- num_total)

```

```{r}

ggplot(df, aes(x=sizes)) + 
  geom_line(aes(y = npos, col = "Number of Positive Sequences")) + 
  geom_line(aes(y = nneg, col = "Number of Negative Sequences")) + 
  geom_line(aes(y = ntotal, col = "Total number of Sequences")) + 
  labs(x = "Min length of the sequences", y="Number of longer sequences")
  

```

## With percentages

```{r}

sizes <- seq(100,400)
perc_total <- NULL
perc_pos <- NULL
perc_neg <- NULL

for (i in sizes){
  ppos <- sum(nchar(PositiveSequences$sequence) > i)/nrow(PositiveSequences)
  pneg <- sum(nchar(NegativeSequences$sequence) > i)/nrow(NegativeSequences)
  ptot <- (sum(nchar(PositiveSequences$sequence) > i) + sum(nchar(NegativeSequences$sequence) > i))/(nrow(PositiveSequences) + nrow(NegativeSequences))
  
  perc_total <- c(perc_total, ptot)
  perc_pos <- c(perc_pos, ppos)
  perc_neg <- c(perc_neg, pneg)
}

df <- data.frame(sizes = sizes, ppos <- perc_pos, pneg <- perc_neg, ptotal <- perc_total)

```

```{r}

ggplot(df, aes(x=sizes)) + 
  geom_line(aes(y = ppos, col = "% of Positive Sequences")) + 
  geom_line(aes(y = pneg, col = "% of Negative Sequences")) + 
  geom_line(aes(y = ptotal, col = "Total % of Sequences")) + 
  labs(x = "Min length of the sequences", y="Percentage of longer sequences")
  

```

And lastly, we will check exactly what length should the sequence have in order to be as balanced as possible.

```{r}

99 + which(abs(num_pos - num_neg) == min(abs(num_pos - num_neg)))

```

So 202 is the desired length of the sequences.