---
title: "Sequences"
author: "Juan A. Botía"
date: "30/07/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

We want to learn how to work with sequences and move this exciting data type into the realm of $R^N$ so we can work in a numeric and multidimensional space and apply all techniques we know how to apply.

# Reading fasta

We have two types of sequences. 
As in this example we are working with the protein sequences of macro domains coming from Figure 4 of the target paper we're working with, we have to use `readAAStringSet()`. If we wanted to read DNA, we would use `readDNAStringSet()`.

```{r}
#Convert csv to fasta
cesev <- read.csv("../Results/badseqs.tsv", stringsAsFactors = FALSE, header=TRUE, sep="\t")
DNA_seq <- DNAStringSet(x=cesev$raw5utr, use.names=TRUE)
```

```{r}
library(Biostrings)
prots = readDNAStringSet("../Results/badseqs.tsv")
prots
```

So we have 63 sequences, with variable width. And all these methods to play with

```{r}
methods(class="DNAStringSet")
```

And if, for example, we wanted a very basic representation of the proteins into $R^N$, we could use 

```{r}
head(alphabetFrequency(DNA_seq))
```

Analogously, we can obtain the consensus matrix (i.e. the level of agreement, in each position of the strings, if we align them from left to right for each codon) as in 

```{r}
cm = consensusMatrix(DNA_seq)
str(cm)
max(width(DNA_seq))
cm[,1:10]
```

# Dealing with biomart promoters

As our sequence file has no FASTA format, we simply read the content and convert it into a `DNAStringSet` format.

```{r}
#pp = read.delim("~/Dropbox/sequences/sources.tar/positivePromoters_geneflank",stringsAsFactors=F,sep=",")
pp = DNA_seq
pplist = NULL
for(i in 1:nrow(pp))
  pplist[[pp[i,2]]] = pp[i,1] 

ppset = DNA_seq#DNAStringSet(pplist)
cmppset = consensusMatrix(ppset)
```

Now we'll create the consensus matrix as a quick way of inspecting the sequences. Before that, remember the DNA alphabet used in this package is a IUPAC code <http://genome.ucsc.edu/goldenPath/help/iupac.html>, as follows

```{r}
IUPAC_CODE_MAP
```

So we can appropriately represent multiple observed alleles in a single sequence. 

```{r}
cmppset = consensusMatrix(ppset)
dim(cmppset)
cmppset[1:5,1:5]

#Do we have multiple alleles in any of the sequences? No
if(max(cmppset[-c(1:4),]) > 0){
  cat("Warning. We were expecting simple sequences. They are not\n")
}else{
  cat("All sequences are single DNA sequences\n")
  #We keep only what is useful
  cmppset = cmppset[1:4,]
}
  
```

Now we can also try interesting representations, as if we use the alphabet frequency
```{r}
alp = alphabetFrequency(ppset)
alp[1:5,1:5]
alp = alp[,1:4]
pcas = prcomp(t(alp))
plot(pcas$rotation[,1],pcas$rotation[,2])
```
 
Now let's do the same looking for differences between disease promoters and normal ones

```{r}
alp = alphabetFrequency(ppset)
alp = alp[,1:4]
np = read.delim("~/Dropbox/sequences/sources.tar/negativePromoters_geneflank",
                stringsAsFactors=F,sep=",")
nplist = NULL
for(i in 1:nrow(np))
  nplist[[np[i,2]]] = np[i,1] 
npset = DNAStringSet(nplist)
nalp = alphabetFrequency(npset)
max(nalp[,5:18])
nalp[1:5,1:5]
nalp = nalp[,1:4]
aalp = rbind(nalp,alp)

pcas = prcomp(t(aalp))
plot(pcas$rotation[,1],pcas$rotation[,2],pch=19,cex=0.3,
     col=c(rep("grey",nrow(nalp)),rep("red",nrow(alp))))
```

Red points correspond to promoters on disease genes. Therefore, we can say there is no appreciable difference, in terms of letter frequency across the 250 characters of the sequence. Red points are all over the place. 

## Trying different k-mers



```{r}
kmers = 6
for(k in kmers){
  alp = oligonucleotideFrequency(ppset,k)
  nalp = oligonucleotideFrequency(npset,k)
  aalp = rbind(nalp,alp)
  pcas = prcomp(t(aalp))
  plot(pcas$rotation[,1],pcas$rotation[,2],pch=19,cex=0.3,
       main=paste0(k,"k-mers"),
     col=c(rep("grey",nrow(nalp)),rep("red",nrow(alp))))
}

```










# References

* This is a good and basic introduction to biostrings <https://kasperdanielhansen.github.io/genbioconductor/html/Biostrings.html>

* This shows how to manage annotation and sequences simultaneously throug GenomicRanges <https://combine-australia.github.io/2017-05-19-bioconductor-melbourne/strings_and_ranges.html>

* This may be useful to start working with the notion of k-mers <https://cran.r-project.org/web/packages/kmer/vignettes/kmer-vignette.html> 

